Market Expectations and Mangerial Turnover

### 2024-05-15 paper.qmd

```{r}
library(dplyr)
library(purrr)
#spells <- read.csv("raw_data/manager_spells_from_manager_urls_mgronly.csv")
spells <- read.csv("raw_data/manager_spells_from_manager_urls.csv")

# Define season_to_date() function here
season_to_date <- function(season) {
  if (is.na(season) | length(gsub("/", "", season)) == 0) {
    return(NA_real_)
  }

  parts <- strsplit(season, "/")[[1]]
  month <- as.numeric(parts[2])
   if (month < 8) {
     year <- paste0("20", substr(parts[1], 1, 2))
   } else {
     year <- paste0("19", substr(parts[1], 1, 2))
   }

  date_str <- paste0("01/08/", year)
  as.Date(paste(date_str, collapse = "/"), "%d/%m/%Y")
}

data_helper <- spells |>
  mutate(row_id=row_number()) |>
  filter(is.na(Finish)&contract_expiry!="-") |>
  select(contract_expiry, Finish,row_id)
data_helper$Finish<-as.Date(map_dbl(data_helper$contract_expiry, season_to_date))
spells |>
  mutate(row_id=row_number()) |>
  left_join(data_helper |> select(-Finish), by="row_id") |>
  mutate(Start = as.Date(Start),
         Finish = as.Date(Finish)) |>
  mutate(days_in_charge = as.numeric(Finish - Start)) ->spells
# find duplicates for name,club,Finish
#spells |> drop_na(Start,Finish) |> group_by(name,club,Finish) |> summarise(n=n()) |> filter(n>1) |> arrange(desc(n))
# remove duplicates
spells |> drop_na(Start,Finish) |> distinct(name,club,Finish,.keep_all = TRUE) |> filter(Finish-Start>0) -> spells

@@

```{r}
#|label: tbl-managertype
#| tbl-cap: "Manager types"
library(kableExtra)
terms_to_remove <- c("1.FC Köln II","Al-Ahli (UAE)","Gimnasia (J)","San Martín (T)",
  "San Martín (SJ)","Racing (Cba)","Juv. Unida (G)","1.FC Köln U19",
  "1.FC Köln U17","1.FC Köln Yth.",
  "1.FC Köln U16"," II","1.FC Köln","Bor. M'gladbach",
  "Hertha BSC","F. Düsseldorf",
  "Velez Mostar","SSV Reutlingen","FV Ravensburg",
  "Fenerbahce","FC Teningen","FV Nimburg","Heart of Midl.",
  "Vetra","Sp.Genzano (PZ)","Honvéd FC","Olympiacos","Honvéd SE","Hungary",
  "ADO Den Haag(A)","Dordrecht'90","Haarlem","AZ Alkmaar","Roda JC","AZ '67",
  "Sparta R.","De Treffers","NEC Nijmegen","FC Oss","VV Gemert","PSV U21")
spells %>%
  mutate(position = str_remove_all(position, paste0("(", paste(terms_to_remove, collapse = "|"), ")"))) -> spells
spells |> mutate(
    staff_position = trimws(tolower(position)),
    staff_position = ifelse(staff_position=="manager","manager",staff_position),
    staff_position = ifelse(staff_position=="assistant manager","assistant manager",staff_position),
    staff_position = ifelse(staff_position=="caretaker manager","caretaker manager",staff_position),
    staff_position = ifelse(str_detect(staff_position,"player-coach"),"player-coach",staff_position),
    staff_position = ifelse(str_detect(staff_position,"director of football"),"director of football",staff_position),
    staff_position = ifelse(str_detect(staff_position,"technical director"),"technical director",staff_position),
    staff_position = ifelse(str_detect(staff_position,"director of professional football"),"director of professional football",staff_position),
    staff_position = ifelse(str_detect(staff_position,"sporting director"),"sporting director",staff_position),
    staff_position = ifelse(str_detect(staff_position,"academy manager"),"academy manager",staff_position),
    staff_position = ifelse(str_detect(staff_position,"advisor"),"advisor",staff_position),
    staff_position = ifelse(str_detect(staff_position,"head of football operations"),"head of football operations",staff_position),
    staff_position = ifelse(str_detect(staff_position,"head of football operations"),"head of football operations",staff_position)) %>%
  {table(.$staff_position[!is.na(.$staff_position)])} |> sort(decreasing = TRUE) |> 
  # make the frequency table into a data frame
  as.data.frame() |>
  # rename the columns
  rename("staff_position" = "Var1", "n" = "Freq") |>
  # order the data frame by the frequency
  arrange(desc(n)) |>
  head(20) |>
  kable(col.names=c("Staff Role","Number")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F)
```

```{r}
#| label: fig-annual-spell-count
#| fig-cap: "Annual count of manager spell ending"
library(ggplot2)
library(dplyr)
# create plot of annual count of spells using a count of finishes per season assume season starts in August
spells |> mutate(season_start = ifelse(month(Start) < 8, year(Start), year(Start) + 1)) |> 
  count(season_start) |> 
  ggplot(aes(x = season_start, y = n)) +
  geom_col(width = 0.5) +
  labs(title = "Annual count of spell ending", x = "Season", y = "Number of spells") +
  theme_minimal() -> spell_count
plotly::ggplotly(spell_count)
```

```{r}
```

>>>>

```{r}
library(dplyr)
library(purrr)
#spells <- read.csv("raw_data/manager_spells_from_manager_urls_mgronly.csv")
spells <- readRDS("raw_data/manager_spells_from_manager_urls.rds")
```

@@

```{r}
#|label: tbl-managertype
#| tbl-cap: "Manager types"
library(kableExtra)
spells %>%
  {table(.$staff_position[!is.na(.$staff_position)])} |> sort(decreasing = TRUE) |> 
  # make the frequency table into a data frame
  as.data.frame() |>
  # rename the columns
  rename("staff_position" = "Var1", "n" = "Freq") |>
  # order the data frame by the frequency
  arrange(desc(n)) |>
  head(20) |>
  kable(col.names=c("Staff Role","Number")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F)
```

```{r}
#| label: fig-annual-spell-count
#| fig-cap: "Annual count of manager spell ending"
library(ggplot2)
library(dplyr)
# create plot of annual count of spells using a count of finishes per season assume season starts in August
spells |> 
  mutate(youth_team = ifelse(str_detect(club, "[A-Z][0-9]{2}|Yth.|Youth"), "Yes", "No")) -> spells

# create a list of european countries
european_countries <- c("albania", "andorra", "austria", "belarus", "belgium", "bosnia-herzegovina", "bulgaria", "croatia", "cyprus", "czech-republic", "denmark", "estonia", "faroe-islands", "finland", "france", "germany", "gibraltar", "greece", "hungary", "iceland", "ireland", "italy", "kosovo", "latvia", "liechtenstein", "lithuania", "luxembourg", "malta", "moldova", "montenegro", "netherlands", "north-macedonia", "northern-ireland", "norway", "poland", "portugal", "romania", "russia", "san-marino", "scotland", "serbia", "slovakia", "slovenia", "spain", "sweden", "switzerland", "turkey", "ukraine", "wales", "england")
# Create a list of north and south american countries
north_american_countries <- c("canada", "mexico", "united-states","USA")
# list of central american countries
central_american_countries <- c("belize", "costa-rica", "el-salvador", "guatemala", "honduras", "nicaragua", "panama")
south_american_countries <- c("argentina", "bolivia", "brazil", "chile", "colombia", "ecuador", "paraguay", "peru", "uruguay", "venezuela")
# Create list of australiasian countries
australasian_countries <- c("australia", "fiji", "new-zealand", "papua-new-guinea", "solomon-islands", "vanuatu")
# Create list of african countries
african_countries <- c("algeria", "angola", "benin", "botswana", "burkina-faso", "burundi", "cameroon", "cape-verde", "central-african-republic", "chad", "comoros", "congo", "djibouti", "egypt", "equatorial-guinea", "eritrea", "eswatini", "ethiopia", "gabon", "gambia", "ghana", "guinea", "guinea-bissau", "ivory-coast", "kenya", "lesotho", "liberia", "libya", "madagascar", "malawi", "mali", "mauritania", "mauritius", "morocco", "mozambique", "namibia", "niger", "nigeria", "rwanda", "sao-tome-and-principe", "senegal", "seychelles", "sierra-leone", "somalia", "south-africa", "south-sudan", "sudan", "tanzania", "togo", "tunisia", "uganda", "zambia", "zimbabwe")
# Create list of asian countries
asian_countries <- c("afghanistan", "armenia", "azerbaijan", "bahrain", "bangladesh", "bhutan", "brunei", "cambodia", "china", "east-timor", "georgia", "india", "indonesia", "iran")

# plot the annual count of spells using a count of finishes per season assume season starts in August and condition on youth_team
spells |> 
  filter(staff_position=="manager") |> 
  mutate(season_start = ifelse(month(Start) < 8, year(Start), year(Start) + 1)) |>
  group_by(season_start,Answer_1) |>
  count(season_start)  |> 
  ggplot(aes(x = season_start, y = n, fill=Answer_1)) +
  geom_col(width = 0.5) +
  labs(title = "", x = "Season", y = "Number of spells") +
  theme_minimal() -> spell_count
plotly::ggplotly(spell_count)
```

### 2024-05-15 paper_new.qmd

```{r}
library(clipr)
unique(spells$Country) |> write_clip()
```

@@

```{r}
library(countries)
list_countries() |> length()
```

>>>>

```{r}
# Install and load the 'countrycode' package if not already installed
if (!requireNamespace("countrycode", quietly = TRUE)) {
  install.packages("countrycode")
}
library(countries)

# Get the list of country names
countries <- list_countries()

# Function to find the country in a text string
find_country_in_text <- function(text, countries) {
  for (country in countries) {
    if (grepl(tolower(country), tolower(text))) {
      return(country)
    }
  }
  return('NA')
}

# Assuming your data is in a data frame named 'df' and the text strings are in a column named 'text'
spells$matched_country <- sapply(spells$Country, function(x) find_country_in_text(x, countries))
# sanity check
table(spells$matched_country)
```

@@

```{r}
library(countries)
# write to to find partial matches within each string in a list of strings
countries <- spells$Country
countries <- tolower(countries)
countries <- gsub(" - | \\. ", ".", countries)
countries <- gsub("not applicable|none|unknown|not found|not available|n/a", NA, countries)
countries <- gsub("i am not|i cannot|i don't|i am just an ai|i'm just an ai", "LLM cannot find answer", countries)
countries <- countries[!is.na(countries)]
country_list <-tolower(list_countries())
sum(countries %in% tolower(list_countries()))
```
